Symbolic Weighted Language Models and\\ Parsing over Infinite Alphabets

### 2021-05-27 main.tex
revise algo best w. notation operators complete label theory
convexity. macro stack

\def\<#1>{\langle #1 \rangle}
\newcommand{\pair}[2]{\langle{#1}, {#2}\rangle}
%\newcommand{\A}{\mathcal{A}}
%\newcommand{\B}{\mathcal{B}}
\newcommand{\D}{\mathbb{D}}
\newcommand{\E}{\mathbb{E}}
%\newcommand{\P}{\mathcal{P}}
\newcommand{\Q}{\mathcal{Q}}
\newcommand{\R}{\mathcal{R}}
\newcommand{\T}{\mathcal{T}}
\newcommand{\W}{\mathbb{W}}
\newcommand{\Semiring}{\mathbb{S}}
\newcommand{\zero}{\mathbb{0}}
\newcommand{\one}{\mathbb{1}}
\newcommand{\dom}{\ensuremath{\mathit{dom}}}

\def\SA{\textsf{sA}\xspace}
\def\WA{\textsf{wA}\xspace}
\def\SWT{\textsf{swT}\xspace}
\def\SWA{\textsf{swA}\xspace}
\def\SWTA{\textsf{swTA}\xspace}
\def\SWVPA{\textsf{sw-VPA}\xspace}
\def\weight{\mathsf{weight}}
\def\wei{\mathsf{w}}
\def\mei{\mathsf{m}}
\def\init{\mathsf{in}}
\def\final{\mathsf{out}}
\newcommand{\call}[1]{\ensuremath #1} %{\ensuremath \langle_{#1}}
\newcommand{\return}[1]{\ensuremath #1} %{\ensuremath {}_{#1}{\rangle}} % $\prescript{}{a}{)}$
\def\Omegai{{\Omega_\mathsf{i}}}
\def\Omegac{{\Omega_\mathsf{c}}}
\def\Omegar{{\Omega_\mathsf{r}}}
\def\Sigmai{{\Sigma_\mathsf{i}}}
\def\Sigmac{{\Sigma_\mathsf{c}}}
\def\Sigmar{{\Sigma_\mathsf{r}}}
\def\Deltai{{\Delta_\mathsf{i}}}
\def\Deltac{{\Delta_\mathsf{c}}}
\def\Deltar{{\Delta_\mathsf{r}}}
\def\Phii{{\Phi_\mathsf{i}}}
\def\Phic{{\Phi_\mathsf{c}}}
\def\Phir{{\Phi_\mathsf{r}}}
\def\Phix{{\Phi_\mathsf{x}}}
\def\Phici{{\Phi_\mathsf{ci}}}
\def\Phicc{{\Phi_\mathsf{cc}}}
\def\Phicr{{\Phi_\mathsf{cr}}}
\def\Phicx{{\Phi_\mathsf{cx}}}
\def\weii{{\wei_\mathsf{i}}}
\def\weic{{\wei_\mathsf{c}}}
\def\weir{{\wei_\mathsf{r}}}
\def\weie{{\wei_\mathsf{e}}}
\def\weiei{{\wei^\mathsf{e}_\mathsf{i}}}
\def\weiec{{\wei^\mathsf{e}_\mathsf{c}}}
\def\weier{{\wei^\mathsf{e}_\mathsf{r}}}
\def\weiex{{\wei^\mathsf{e}_\mathsf{x}}}
\newcommand{\config}[2]{\ensuremath \genfrac{[}{]}{0pt}{}{#1}{#2}} 
\newcommand{\opluseq}{\ensuremath\mathrel{\oplus}=}
\newcommand{\ioi}[1]{\mathsf{ioi}({#1})}
\newcommand{\rank}{\mathsf{rk}}
\newcommand{\lin}{\mathsf{lin}}

@@

Formally, the transitions of the automaton~$A$ are defined
in term of %a weight value computed by 
an intermediate function $\weight_A$, like in Section~\ref{sec:SWT}.
In the case of a pushdown automaton, a configuration is composed 
of a state $q \in Q$ and a stack content $\gamma \in \Gamma^*$, where $\Gamma = \Omegac \times P$.
Hence, $\weight_A$ is a function from 
$[Q \times \Gamma^*] \times \Omega^* \times [Q \times \Gamma^*]$ into~$\Semiring$.
%
\begin{align}
%\begin{array}[t]{rcl}
\weight_A\bigl(\config{q}{{\call{c}}\, p\cdot\gamma}, a\, u, \config{q'}{\gamma'}\bigr) & =  
 {\displaystyle\bigoplus_{q'' \in Q}} \weii(q, c, p, a, q'') 
 \otimes \weight_A\bigl(\config{q''}{\gamma}, u, \config{q'}{\gamma'}\bigr)\nonumber\\
\weight_A\bigl(\config{q}{\bot}, a\, u, \config{q'}{\gamma'}\bigr) & =  
 {\displaystyle\bigoplus_{q'' \in Q}} \weiei(q, a, q'') 
 \otimes \weight_A\bigl(\config{q''}{\bot}, u, \config{q'}{\gamma'}\bigr)\nonumber\\
%
\weight_A\bigl(\config{q}{{\call{c}}\, p\cdot\gamma}, {\call{c}'}\, u, \config{q'}{\gamma'}\bigr) & =  
 {\displaystyle\bigoplus_{\begin{array}{c}
                          \scriptstyle q'' \in Q\\[-2pt]
                          \scriptstyle p \in P
                          \end{array}}}
 \weic\bigl(q, {\call{c}}, p, {\call{c}'}, p', q''\bigr) 
 \otimes \weight_A\bigl(\config{q''}{{\call{c}'}\, p' \cdot {\call{c}}\, p\cdot \gamma}, u, \config{q'}{\gamma'}\bigr)\nonumber\\[1mm]
%
\weight_A\bigl(\config{q}{\bot}, {\call{c}}\, u, \config{q'}{\gamma'}\bigr) & =  
 {\displaystyle\bigoplus_{q'' \in Q}} \weiec(q, {\call{c}}, p, q'') 
 \otimes \weight_A\bigl(\config{q''}{{\call{c}}\, p}, u, \config{q'}{\gamma'}\bigr)\nonumber\\
%
\weight_A\bigl(\config{q}{{\call{c}}\, p\cdot \gamma}, {\return{r}}\, u, \config{q'}{\gamma'}\bigr) & =  
 {\displaystyle\bigoplus_{q'' \in Q}} 
 \weir\bigl(q, {\call{c}}, p, {\return{r}}, q''\bigr) 
 \otimes \weight_A\bigl(\config{q''}{\gamma}, u, \config{q'}{\gamma'}\bigr)\nonumber\\
%
\weight_A\bigl(\config{q}{\bot}, {\return{r}}\, u, \config{q'}{\gamma'}\bigr) & =  
 {\displaystyle\bigoplus_{q'' \in Q}} \weier(q, {\return{r}}, q'') 
 \otimes \weight_A\bigl(\config{q''}{\bot}, u, \config{q'}{\gamma'}\bigr)
%\end{array}
\label{eq:SWVPA-weight}
\end{align}
%
where $\bot$ denotes the empty stack and ${\call{c}}\, p\cdot \gamma$ 
denotes a stack where the pair made of ${\call{c}} \in \Omegac$ and $p \in P$ is the top symbol 
and $\gamma$ is the rest of stack.

@@

%
\noindent
Let $\top$ be a fresh stack symbol which does not belong to $\Gamma$,
and let $b_\top : Q \times P \times Q \to \Phic$ be such that,
for every two states $q, q' \in Q$ 
and stack symbol $p \in P$: %and $\sigma \in \{ \bot, \top \}$, 
\begin{equation}\label{eq:btop}
  b_\top(q, p, q') : c \mapsto \bigoplus_{s\in \Omega^*} 
  \textstyle
  \weight_A\bigl(\config{q}{c\,p\cdot \top }, s, \config{q'}{c\, p\cdot \top}\bigr). 
\end{equation}
%
Intuitively, the function defined in \eqref{eq:btop}
associateds to $c \in \Omegac$ 
the minimum weight of a computation of $A$
starting in state $q$ with a stack 
$c p \cdot \gamma \in \Gamma^+$ 
and ending in state $q'$ with the same stack,
such that the computation does pop 
the pair made of $c$ and $p$ at the top of this stack,
but may read these symbols.
Moreover, $A$ may push another pair $\< c', p'>$ %call symbols 
on the top of $c p \cdot \gamma$,
following the the third case of 
in the definition \eqref{eq:SWVPA-weight} of $\weight_A$,
and may pop $\< c', p'>$ later, following the fifth case of \eqref{eq:SWVPA-weight} (return symbol). 
%However, it cannot apply one of the two last cases (return symbol and empty stack)
%when the current stack is $\gamma$.
%pop symbols in $\gamma$.
% Note that having a stack reduced to such a symbol makes impossible the application of the 
% two last cases in the definition of $\weight_A$ (return symbol and empty stack). 
% However, it is possible to apply the two first cases 
% (internal symbol or call symbol, with a push on the top of $\top$).

@@

\begin{figure}
For all $q_0, q_3 \in Q$, %$p \in P$,
\[
\begin{array}{lcl}
%\multicolumn{3}{l}{\mathrm{For~all~} q_0, q_3 \in Q, p \in P,}\\
d_\top(q_1, p, q_3) & \opluseq &
  d_\top(q_1, p, q_2) \otimes 
  \displaystyle\bigoplus_{a \in \Omegai} \weii(q_2, p, q_3)_a\\
\multicolumn{3}{l}{\quad
\mathrm{where~} \weii(q_2, p, q_3)_a \in \Phic
\mathrm{~is~the~partial~application~}
x_c \mapsto \weii(q_2, x_c, p, a, q_3)}\\
%
d_\bot(q_1, p, q_3) & \opluseq &
  d_\bot(q_1, q_2) \otimes 
  \displaystyle\bigoplus_{a \in \Omegai} \weie(q_2, a, q_3)\\
%     
d_\top(q_0, p, q_3) & \opluseq &
  \displaystyle\bigoplus_{{\call{c}'} \in \Omegac} 
  \displaystyle\bigoplus_{{\return{r}} \in \Omegar} 
  \weic(q_0, p, p', q_1)_{c'} \otimes 
  d_\top(q_1, p', q_2) \otimes
  \weir(q_2, c', p', r, q_3)\\
\multicolumn{3}{l}{\quad
\mathrm{where~} \weic(q_0, p, p', q_1)_{c'} \in \Phic
\mathrm{~is~the~partial~application~}
x_c \mapsto \weic(q_0, p, x_c, c', p', q_1)}\\[2pt]
%
d_\bot(q_0, q_3) & \opluseq &
  \displaystyle\bigoplus_{{\call{c}} \in \Omegac} 
  \displaystyle\bigoplus_{{\return{r}} \in \Omegar} 
  \weiec(q_0, c, p, q_1) \otimes 
   d_\top(q_1, p, q_2)(c) \otimes
  \weir(q_2, c, p, r, q_3)\\
%
d_\bot(q_1, q_3) & \opluseq &
  d_\bot(q_1, q_2) \otimes 
  \displaystyle\bigoplus_{\return{r} \in \Omegar} \weier(q_2, {\return{r}}, q_3)\\
%
d_\top(q_1, p, q_3) & \opluseq & 
  d_\top(q_1, p, q_2) \otimes d_\top(q_2, p, q_3), 
  \mathrm{if~} \< q_2, \top, q_3> \notin P \\
%
d_\bot(q_1, q_3) & \opluseq & 
  d_\bot(q_1, q_2) \otimes d_\bot(q_2, q_3), \mathrm{if~} \< q_2, \bot, q_3> \notin P \\
\end{array}
\]
%$\weii(q_2, p, q_3)_a \in \Phic$, for $a \in \Omegai$,
%is the partial application
%$x_c \mapsto \weii(q_2, x_c, p, a, q_3)$.
%
%$\weic(q_0, p, p', q_1)_{c'} \in \Phic$, for $c' \in \Omegac$,
%is the partial application
%$x_c \mapsto \weic(q_0, p, x_c, c', p', q_1)$.
\caption{Update $d_\bot$ with $\<q_1, q_2>$ or $d_\top$ with $\< q_1, p, q_2>$.} 
\label{fig:best-update}
\end{figure}

@@

%\medskip
\noindent
In order to obtain effectively a witness 
(word of $\Semiring^*$ with computation of $A$ of minimal weight), 
we require an additional property the of weight functions.
%
\begin{definition}
Let $\Omega$ be an alphabet and $\Semiring$ a complete semiring.
A function $\phi$ from $\Omega^n$ into $\Semiring$
is called $k$-\emph{convex} for a natural number $k$ iff 
$\mathit{card}\{ \mathbf{a} \in \Omega^n \mid \phi(\mathbf{a}) = 
                 \displaystyle\bigoplus_{\mathbf{p} \in \Omega^n} \phi(\mathbf{p}) \} \leq k$.
\end{definition}
A label theory is $k$-convex if all its functions are $k$-convex.

\begin{proposition}\label{th:best-search}
For a \SWVPA $A$ 
over $\Omega$, 
$\Semiring$ commutative, bounded, total and complete, %semiring 
and $\bar\Phi$ $k$-convex effective, % label theory,
one can construct in PTIME a word $s \in \Omega^*$ 
such that $A(s)$ is minimal \wrt the natural ordering for $\Semiring$. 
\end{proposition}

>>>>

\def\<#1>{\langle #1 \rangle}
\newcommand{\pair}[2]{\langle{#1}, {#2}\rangle}

%\newcommand{\config}[2]{\ensuremath \genfrac{[}{]}{0pt}{}{#1}{#2}} 
\newcommand{\config}[2]{\ensuremath{#1}[{#2}]} 
\newcommand{\configup}[2]{\ensuremath{#1}\left[\begin{array}{c} #2 \end{array}\right]}
\def\stacksep{\cdot}
\def\stackup{\\}

\newcommand{\opluseq}{\ensuremath\mathrel{\oplus}=}

% \bigominus
\DeclareFontFamily{U}{mathx}{\hyphenchar\font45}
\DeclareFontShape{U}{mathx}{m}{n}{
<-6> mathx5 <6-7> mathx6 <7-8> matha7
<8-9> mathx8 <9-10> mathx9
<10-12> mathx10 <12-> mathx12
}{}
\DeclareSymbolFont{mathx}{U}{mathx}{m}{n}
\DeclareMathSymbol{\bigominus}{\mathop}{mathx}{"C1}

%\newcommand{\A}{\mathcal{A}}
%\newcommand{\B}{\mathcal{B}}
\newcommand{\D}{\mathbb{D}}
\newcommand{\E}{\mathbb{E}}
%\newcommand{\P}{\mathcal{P}}
\newcommand{\Q}{\mathcal{Q}}
\newcommand{\R}{\mathcal{R}}
\newcommand{\T}{\mathcal{T}}
\newcommand{\W}{\mathbb{W}}
\newcommand{\Semiring}{\mathbb{S}}
\newcommand{\zero}{\mathbb{0}}
\newcommand{\one}{\mathbb{1}}
\newcommand{\dom}{\ensuremath{\mathit{dom}}}

\def\SA{\textsf{sA}\xspace}
\def\WA{\textsf{wA}\xspace}
\def\SWT{\textsf{swT}\xspace}
\def\SWA{\textsf{swA}\xspace}
\def\SWTA{\textsf{swTA}\xspace}
\def\SWVPA{\textsf{sw-VPA}\xspace}
\def\weight{\mathsf{weight}}
\def\wei{\mathsf{w}}
\def\mei{\mathsf{m}}
\def\init{\mathsf{in}}
\def\final{\mathsf{out}}
\newcommand{\call}[1]{\ensuremath #1} %{\ensuremath \langle_{#1}}
\newcommand{\return}[1]{\ensuremath #1} %{\ensuremath {}_{#1}{\rangle}} % $\prescript{}{a}{)}$
\def\Omegai{{\Omega_\mathsf{i}}}
\def\Omegac{{\Omega_\mathsf{c}}}
\def\Omegar{{\Omega_\mathsf{r}}}
\def\Sigmai{{\Sigma_\mathsf{i}}}
\def\Sigmac{{\Sigma_\mathsf{c}}}
\def\Sigmar{{\Sigma_\mathsf{r}}}
\def\Deltai{{\Delta_\mathsf{i}}}
\def\Deltac{{\Delta_\mathsf{c}}}
\def\Deltar{{\Delta_\mathsf{r}}}
\def\Phii{{\Phi_\mathsf{i}}}
\def\Phic{{\Phi_\mathsf{c}}}
\def\Phir{{\Phi_\mathsf{r}}}
\def\Phix{{\Phi_\mathsf{x}}}
\def\Phici{{\Phi_\mathsf{ci}}}
\def\Phicc{{\Phi_\mathsf{cc}}}
\def\Phicr{{\Phi_\mathsf{cr}}}
\def\Phicx{{\Phi_\mathsf{cx}}}
\def\weii{{\wei_\mathsf{i}}}
\def\weic{{\wei_\mathsf{c}}}
\def\weir{{\wei_\mathsf{r}}}
\def\weie{{\wei_\mathsf{e}}}
\def\weiei{{\wei^\mathsf{e}_\mathsf{i}}}
\def\weiec{{\wei^\mathsf{e}_\mathsf{c}}}
\def\weier{{\wei^\mathsf{e}_\mathsf{r}}}
\def\weiex{{\wei^\mathsf{e}_\mathsf{x}}}
\newcommand{\ioi}[1]{\mathsf{ioi}({#1})}
\newcommand{\rank}{\mathsf{rk}}
\newcommand{\lin}{\mathsf{lin}}

@@

Formally, the transitions of the automaton~$A$ are defined
in term of %a weight value computed by 
an intermediate function $\weight_A$, like in Section~\ref{sec:SWT}.
In the case of a pushdown automaton, a configuration, denoted by $q[\gamma]$ 
is composed of a state $q \in Q$ 
and a stack content $\gamma \in \Gamma^*$, 
where $\Gamma = \Omegac \times P$.
Hence, $\weight_A$ is a function from 
$[Q \times \Gamma^*] \times \Omega^* \times [Q \times \Gamma^*]$ into~$\Semiring$.
%
\begin{align}
%\begin{array}[t]{rcl}
\weight_A\bigl(\configup{q}{\<{\call{c}}, p>\stackup\gamma}, a\, u, 
               \config{q'}{\gamma'}\bigr) & =  
 {\displaystyle\bigoplus_{q'' \in Q}} \weii(q, c, p, a, q'') 
  \otimes \weight_A\bigl(\configup{q''}{\<{\call{c}}, p> \stackup \gamma}, u, 
                         \config{q'}{\gamma'}\bigr)\nonumber\\
 %
\weight_A\bigl(\config{q}{\bot}, a\, u, 
               \config{q'}{\gamma'}\bigr) & =  
  {\displaystyle\bigoplus_{q'' \in Q}} \weiei(q, a, q'') 
   \otimes \weight_A\bigl(\config{q''}{\bot}, u, \config{q'}{\gamma'}\bigr)\nonumber\\
%
\weight_A\bigl(\configup{q}{\<{\call{c}}, p>\stackup\gamma}, {\call{c}'}\, u, 
               \config{q'}{\gamma'}\bigr) & =  
 {\displaystyle\bigoplus_{\begin{array}{c}
                          \scriptstyle q'' \in Q\\[-2pt]
                          \scriptstyle p \in P
                          \end{array}}}
 \weic\bigl(q, {\call{c}}, p, {\call{c}'}, p', q''\bigr) 
 \otimes \weight_A\bigl(\configup{q''}{\<{\call{c}'}, p'>\stackup \<{\call{c}}, p>\stackup \gamma}, u, 
                        \config{q'}{\gamma'}\bigr)\nonumber\\[1mm]
%
\weight_A\bigl(\config{q}{\bot}, {\call{c}}\, u, 
               \config{q'}{\gamma'}\bigr) & =  
 {\displaystyle\bigoplus_{q'' \in Q}} \weiec(q, {\call{c}}, p, q'') 
  \otimes \weight_A\bigl(\config{q''}{{\call{c}}\, p}, u, 
                         \config{q'}{\gamma'}\bigr)\nonumber\\
%
\weight_A\bigl(\configup{q}{\<{\call{c}}, p>\stackup \gamma}, {\return{r}}\, u, 
               \config{q'}{\gamma'}\bigr) & =  
 {\displaystyle\bigoplus_{q'' \in Q}} 
  \weir\bigl(q, {\call{c}}, p, {\return{r}}, q''\bigr) 
  \otimes \weight_A\bigl(\config{q''}{\gamma}, u, 
                         \config{q'}{\gamma'}\bigr)\nonumber\\
%
\weight_A\bigl(\config{q}{\bot}, {\return{r}}\, u, 
               \config{q'}{\gamma'}\bigr) & =  
 {\displaystyle\bigoplus_{q'' \in Q}} \weier(q, {\return{r}}, q'') 
  \otimes \weight_A\bigl(\config{q''}{\bot}, u, 
                         \config{q'}{\gamma'}\bigr)
%\end{array}
\label{eq:SWVPA-weight}
\end{align}
%
where $\bot$ denotes the empty stack.
% and ${\call{c}}\, p\stacksep \gamma$ 
%denotes a stack where the pair made of ${\call{c}} \in \Omegac$ and $p \in P$ is the top symbol 
%and $\gamma$ is the rest of stack.

@@

%
\noindent
Let $\top$ be a fresh stack symbol which does not belong to $\Gamma$,
and let $b_\top : Q \times P \times Q \to \Phic$ be such that,
for every two states $q, q' \in Q$ 
and stack symbol $p \in P$: %and $\sigma \in \{ \bot, \top \}$, 
\begin{equation}\label{eq:btop}
  b_\top(q, p, q') : c \mapsto \bigoplus_{s\in \Omega^*} 
  \textstyle
  \weight_A\bigl(\configup{q}{\< c, p> \stackup \top }, s, \configup{q'}{\<c, p> \stackup \top}\bigr). 
\end{equation}
%
Intuitively, the function defined in \eqref{eq:btop}
associateds to $c \in \Omegac$ 
the minimum weight of a computation of $A$
starting in state $q$ with a stack 
$c p \cdot \gamma \in \Gamma^+$ 
and ending in state $q'$ with the same stack,
such that the computation does pop 
the pair made of $c$ and $p$ at the top of this stack,
but may read these symbols.
Moreover, $A$ may push another pair $\< c', p'>$ %call symbols 
on the top of $c p \cdot \gamma$,
following the the third case of 
in the definition \eqref{eq:SWVPA-weight} of $\weight_A$,
and may pop $\< c', p'>$ later, following the fifth case of \eqref{eq:SWVPA-weight} (return symbol). 
%However, it cannot apply one of the two last cases (return symbol and empty stack)
%when the current stack is $\gamma$.
%pop symbols in $\gamma$.
% Note that having a stack reduced to such a symbol makes impossible the application of the 
% two last cases in the definition of $\weight_A$ (return symbol and empty stack). 
% However, it is possible to apply the two first cases 
% (internal symbol or call symbol, with a push on the top of $\top$).

@@

\begin{figure}
For all $q_0, q_3 \in Q$, %$p \in P$,
\[
\begin{array}{lcl}
%\multicolumn{3}{l}{\mathrm{For~all~} q_0, q_3 \in Q, p \in P,}\\
d_\top(q_1, p, q_3) & \opluseq &
  d_\top(q_1, p, q_2) \otimes 
  \displaystyle\bigoplus_{\Omegai} \weii(q_2, p, q_3)\\
%\multicolumn{3}{l}{\quad
%\mathrm{where~} \weii(q_2, p, q_3)_a \in \Phic
%\mathrm{~is~the~partial~application~}
%x_c \mapsto \weii(q_2, x_c, p, a, q_3)}\\
%
d_\bot(q_1, p, q_3) & \opluseq &
  d_\bot(q_1, q_2) \otimes 
  \displaystyle\bigoplus_{\Omegai} \weiei(q_2, q_3)\\
%     
d_\top(q_0, p, q_3) & \opluseq &
  {\displaystyle\bigoplus_{\Omegac}}^2 
  \bigl[ \bigl( \weic(q_0, p, p', q_1) \otimes_2 
  d_\top(q_1, p', q_2) \bigr) \otimes_2
  {\displaystyle\bigoplus_{\Omegar}} \weir(q_2, p', q_3) \bigr]\\
%\multicolumn{3}{l}{\quad
%\mathrm{where~} \weic(q_0, p, p', q_1)_{c'} \in \Phic
%\mathrm{~is~the~partial~application~}
%x_c \mapsto \weic(q_0, p, x_c, c', p', q_1)}\\[2pt]
%
d_\bot(q_0, q_3) & \opluseq &
  {\displaystyle\bigoplus_{\Omegac}}
  \bigl(
  \weiec(q_0, p, q_1) \otimes 
   d_\top(q_1, p, q_2) \otimes
  \displaystyle\bigoplus_{\Omegar} \weir(q_2, p, q_3)\bigr)\\
%
d_\bot(q_1, q_3) & \opluseq &
  d_\bot(q_1, q_2) \otimes 
  \displaystyle\bigoplus_{\Omegar} \weier(q_2, q_3)\\
%
d_\top(q_1, p, q_3) & \opluseq & 
  d_\top(q_1, p, q_2) \otimes d_\top(q_2, p, q_3), 
  \mathrm{if~} \< q_2, \top, q_3> \notin P \\
%
d_\bot(q_1, q_3) & \opluseq & 
  d_\bot(q_1, q_2) \otimes d_\bot(q_2, q_3), \mathrm{if~} \< q_2, \bot, q_3> \notin P \\
\end{array}
\]
%$\weii(q_2, p, q_3)_a \in \Phic$, for $a \in \Omegai$,
%is the partial application
%$x_c \mapsto \weii(q_2, x_c, p, a, q_3)$.
%
%$\weic(q_0, p, p', q_1)_{c'} \in \Phic$, for $c' \in \Omegac$,
%is the partial application
%$x_c \mapsto \weic(q_0, p, x_c, c', p', q_1)$.
\caption{Update $d_\bot$ with $\<q_1, q_2>$ or $d_\top$ with $\< q_1, p, q_2>$.} 
\label{fig:best-update}
\end{figure}

@@

%\medskip
\noindent
In order to obtain effectively a witness 
(word of $\Omega^*$ with a computation of $A$ of minimal weight), 
we require the additional property of convexity of weight functions.

\begin{proposition}\label{th:best-search}
For a \SWVPA $A$ 
over $\Omega$, 
$\Semiring$ commutative, bounded, total and complete, %semiring 
and $\bar\Phi$ effective and $k$-convex, % label theory,
one can construct in PTIME a word $s \in \Omega^*$ 
such that $A(s)$ is minimal \wrt the natural ordering for $\Semiring$. 
\end{proposition}
